### 2022-03-22 paper/output/document/thesis_0_abstract.tex
i hate don knuth

In this thesis, I present an implementation of automated Kantian
ethics that is faithful to the Kantian philosophical tradition. Given an appropriately represented 
action and minimal factual background, my system can judge a potential action as morally obligatory, permissible, or prohibited.
To accomplish this, I formalize Kant's categorical imperative, or moral rule,
in Carmo and Jones's Dyadic Deontic Logic, implement this formalization 
in the Isabelle/HOL theorem prover, and develop a testing framework to evaluate how well 
my implementation coheres with expected properties of Kantian ethics, as established in the literature. 
I also use my system to derive philosophically sophisticated and nuanced solutions to two central ethical 
dilemmas in Kantian literature: the permissibility of lying in the context of a joke and to a murderer 
asking about the location of their intended victim. Finally, I examine the philosophical implications 
of my system, exploring its potential to guide AI agents, academic philosophers, 
and everyday reasoners as they navigate ethical dilemmas. Ultimately, this work serves 
as an early proof-of-concept for philosophically mature AI agents and is one step towards the development 
of responsible, trustworthy artificial intelligence.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/Desktop/cs91r/paper/thesis_0_abstract.thy%:%
%:%19=6%:%
%:%20=7%:%
%:%21=8%:%
%:%22=9%:%
%:%23=10%:%
%:%24=11%:%
%:%25=12%:%
%:%26=13%:%
%:%27=14%:%
%:%28=15%:%
%:%29=16%:%
%:%30=17%:%
%:%31=18%:%
%:%32=19%:%
%:%33=20%:%
%:%34=21%:%
%:%35=22%:%
%:%36=23%:%
%:%37=24%:%
%:%38=25%:%
%:%39=26%:%
%:%40=27%:%
%:%41=28%:%
%:%42=29%:%
%:%43=30%:%
%:%44=31%:%
%:%45=32%:%

>>>>

In this thesis, I present an implementation of automated Kantian
ethics that is faithful to the Kantian philosophical tradition. Given minimal factual background, my 
system can judge an agent's potential action as morally obligatory, permissible, or prohibited.
To accomplish this, I formalize Kant's categorical imperative, or moral rule,
in deontic logic, implement this formalization 
in the Isabelle/HOL theorem prover, and develop a testing framework to evaluate how well 
my implementation coheres with expected properties of Kantian ethics, as established in the literature. 
This testing framework demonstrates that my system outperforms two other potential implementations of 
automated Kantian ethics. I also use my system to derive philosophically sophisticated and nuanced 
solutions to two central controversies in Kantian literature: the permissibility of lying in the 
context of a joke and to a murderer asking about the location of their intended victim. Finally, I 
examine my system's philosophical implications, demonstrating that it can not only guide AI agents, but it can
also help academic philosophers make philosophical progress and augment the everyday ethical reasoning
that we all perform as we navigate the world. I contribute a working proof-of-concept implementation
of automated Kantian ethics capable of performing philosophical reasoning more mature than anything previously
automated. My work serves as one step towards the development of responsible, trustworthy artifical intelligence.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/Desktop/cs91r/paper/thesis_0_abstract.thy%:%
%:%19=6%:%
%:%20=7%:%
%:%21=8%:%
%:%22=9%:%
%:%23=10%:%
%:%24=11%:%
%:%25=12%:%
%:%26=13%:%
%:%27=14%:%
%:%28=15%:%
%:%29=16%:%
%:%30=17%:%
%:%31=18%:%
%:%32=19%:%
%:%33=20%:%
%:%34=21%:%
%:%35=22%:%
%:%36=23%:%
%:%37=24%:%
%:%38=25%:%
%:%39=26%:%
%:%40=27%:%
%:%41=28%:%
%:%42=29%:%
%:%43=30%:%
%:%44=31%:%
%:%45=32%:%
%:%46=33%:%
%:%47=34%:%

### 2022-03-22 paper/output/document/thesis_1_intro.tex
i hate don knuth

My implementation of automated Kantian ethics formalizes Kant's moral rule
in deontic logic, a modal logic that can express obligation, or morally binding requirements. Most modal 
logics include a necessitation operator, denoted as $\Box$, where
$\Box p$ is true at world $w$ if $p$ is true at all worlds that neighbor $w$ \citep{cresswell}. Such 
logics  also contain the possibility operator $\diamond$, where $\diamond \, p \longleftrightarrow \neg (\Box (\neg p))$ 
and operators of propositional logic like $\neg, \wedge, \vee, \rightarrow$. Standard deontic logic (SDL) 
replaces $\Box$ with the obligation operator $O$, where $O \, p$ is true at $w$ 
if $p$ is true at all morally perfect versions of $w$ \citep{sep-logic-deontic}. I use a more sophisticated
version of DDL called Dyadic Deontic Logic, in which
the dyadic obligation operator $O\{A \vert B\}$ represents the sentence ``A is obligated in the context B.'' 
The introduction of context allows DDL to express more nuanced reasoning and certain paradoxes that plague
SDL.

@@

A machine that can evaluate the moral status of a maxim can not only help machines better reason about ethics, 
but it can also help philosophers better study philosophy. I argue for ``computational ethics,'' or the use of computational tools to 
make philosophical progress, analogous to computational biology. 
I demonstrate the potential of computational ethics by presenting a 
philosophical insight about which kinds of actions are appropriate for ethical consideration that I 
discovered using my system. The process of building and interacting with a computer that can reason 
about ethics helped me, a human philosopher, arrive at a philosophical conclusion that has implications for practical
reason and philosophy of doubt. Thus, my system can be used in three distinct ways. First, my system can help
automated agents navigate the world, which I will refer to as automated ethics or machine ethics interchangeably. Second, 
my system help human philosophers reason about philosophy, which I call computational ethics. Third, as 
I discuss in Section \ref{ordinarypeople}, computational ethics can help not only professional philosophers,
but can also augment the everyday ethical reasoning that we all perform as we navigate the world.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/Desktop/cs91r/paper/thesis_1_intro.thy%:%
%:%24=6%:%
%:%36=8%:%
%:%37=9%:%
%:%38=10%:%
%:%39=11%:%
%:%40=12%:%
%:%41=13%:%
%:%42=14%:%
%:%43=15%:%
%:%44=16%:%
%:%45=17%:%
%:%46=18%:%
%:%47=19%:%
%:%48=20%:%
%:%49=21%:%
%:%50=22%:%
%:%51=23%:%
%:%52=24%:%
%:%53=25%:%
%:%54=26%:%
%:%55=27%:%
%:%56=28%:%
%:%57=29%:%
%:%58=30%:%
%:%59=31%:%
%:%60=32%:%
%:%61=33%:%
%:%62=34%:%
%:%63=35%:%
%:%64=36%:%
%:%65=37%:%
%:%66=38%:%
%:%67=39%:%
%:%68=40%:%
%:%69=41%:%
%:%70=42%:%
%:%71=43%:%
%:%72=44%:%
%:%73=45%:%
%:%74=46%:%
%:%75=47%:%
%:%76=48%:%
%:%77=49%:%
%:%78=50%:%
%:%79=51%:%
%:%80=52%:%
%:%81=53%:%
%:%82=54%:%
%:%83=55%:%
%:%84=56%:%
%:%85=57%:%
%:%86=58%:%
%:%87=59%:%
%:%88=60%:%
%:%89=61%:%
%:%90=62%:%
%:%91=63%:%
%:%92=64%:%
%:%93=65%:%
%:%94=66%:%
%:%95=67%:%
%:%96=68%:%
%:%97=69%:%
%:%98=70%:%
%:%99=71%:%
%:%100=72%:%
%:%101=73%:%
%:%102=74%:%
%:%103=75%:%
%:%104=76%:%
%:%105=77%:%
%:%106=78%:%
%:%107=79%:%
%:%108=80%:%
%:%109=81%:%
%:%110=82%:%
%:%111=83%:%
%:%112=84%:%
%:%113=85%:%
%:%114=86%:%
%:%115=87%:%
%:%116=88%:%
%:%120=91%:%
%:%121=92%:%
%:%122=93%:%
%:%123=94%:%
%:%124=95%:%
%:%125=96%:%
%:%126=97%:%
%:%127=98%:%
%:%128=99%:%
%:%129=100%:%
%:%130=101%:%
%:%131=102%:%
%:%132=103%:%
%:%133=104%:%
%:%134=105%:%
%:%135=106%:%
%:%136=107%:%
%:%137=108%:%
%:%138=109%:%
%:%139=110%:%
%:%140=111%:%
%:%141=112%:%
%:%142=113%:%
%:%143=114%:%
%:%144=115%:%
%:%145=116%:%
%:%149=119%:%
%:%150=120%:%
%:%151=121%:%
%:%152=122%:%
%:%153=123%:%
%:%154=124%:%
%:%155=125%:%
%:%156=126%:%
%:%157=127%:%
%:%158=128%:%
%:%159=129%:%
%:%160=130%:%
%:%161=131%:%
%:%162=132%:%
%:%163=133%:%
%:%164=134%:%
%:%165=135%:%
%:%166=136%:%
%:%167=137%:%
%:%168=138%:%
%:%169=139%:%
%:%170=140%:%
%:%171=141%:%
%:%172=142%:%
%:%173=143%:%
%:%174=144%:%
%:%175=145%:%
%:%176=146%:%
%:%177=147%:%
%:%178=148%:%
%:%179=149%:%
%:%180=150%:%
%:%181=151%:%
%:%182=152%:%
%:%183=153%:%
%:%184=154%:%
%:%185=155%:%
%:%186=156%:%
%:%187=157%:%
%:%188=158%:%
%:%189=159%:%
%:%190=160%:%
%:%191=161%:%
%:%192=162%:%
%:%193=163%:%
%:%194=164%:%
%:%195=165%:%
%:%196=166%:%
%:%197=167%:%
%:%198=168%:%
%:%199=169%:%
%:%200=170%:%
%:%201=171%:%
%:%202=172%:%
%:%203=173%:%
%:%204=174%:%
%:%205=175%:%
%:%206=176%:%
%:%207=177%:%
%:%208=178%:%
%:%209=179%:%
%:%210=180%:%
%:%211=181%:%
%:%212=182%:%
%:%213=183%:%
%:%214=184%:%
%:%215=185%:%
%:%216=186%:%
%:%217=187%:%

>>>>

My implementation of automated Kantian ethics formalizes Kant's moral rule
in deontic logic, a modal logic that can express obligation, or morally binding requirements. Most modal 
logics include a necessitation operator, denoted as $\Box$, where
$\Box p$ is true at world $w$ if $p$ is true at all worlds that neighbor $w$ \citep{cresswell}. Intuitively, 
$\Box p$ indicates that $p$ is a necessary truth, like $p$ or $\neg p$. Such 
logics also contain the $\diamond$ operator, which represents possible truths, and operators of 
propositional logic like $\neg, \wedge, \vee, \rightarrow$. Deontic logics
replace $\Box$ with the obligation operator $O$, where $O \, p$ is true at $w$ 
if $p$ is true at all morally perfect versions of $w$ \citep{sep-logic-deontic}. A necessary 
proposition must be true, while an obligatory proposition must be true in order for a world to be morally good.
For example, in order for a world to be morally perfect, if giving to charity is morally obligatory, then
the statement ``Sara gives to charity'' must be true at that world. I use a sophisticated deontic logic 
called Dyadic Deontic Logic, in which the dyadic obligation operator $O\{A \vert B\}$ represents the 
sentence ``A is obligated in the context B.'' This operator expresses the nuanced idea
that certain acts are morally required in certain situations, but not in others.

@@

A machine that can evaluate the moral status of an action can not only help machines better reason about ethics, 
but it can also help philosophers better study philosophy. I argue for ``computational ethics,'' or the use of computational tools to 
make philosophical progress, analogous to computational biology. 
I demonstrate the potential of computational ethics by presenting a 
philosophical insight about which kinds of actions are appropriate for ethical consideration that I 
discovered using my system. The process of building and interacting with a computer that can reason 
about ethics helped me, a human philosopher, arrive at a philosophical conclusion that has implications for practical
reason and philosophy of doubt. Thus, my system can be used in three distinct ways. First, my system can help
automated agents navigate the world, which I will refer to as automated ethics or machine ethics interchangeably. Second, 
my system help human philosophers reason about philosophy, which I call computational ethics. Third, as 
I discuss in Section \ref{ordinarypeople}, computational ethics can help not only professional philosophers,
but can also augment the everyday ethical reasoning that we all perform as we navigate the world.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/Desktop/cs91r/paper/thesis_1_intro.thy%:%
%:%24=6%:%
%:%36=8%:%
%:%37=9%:%
%:%38=10%:%
%:%39=11%:%
%:%40=12%:%
%:%41=13%:%
%:%42=14%:%
%:%43=15%:%
%:%44=16%:%
%:%45=17%:%
%:%46=18%:%
%:%47=19%:%
%:%48=20%:%
%:%49=21%:%
%:%50=22%:%
%:%51=23%:%
%:%52=24%:%
%:%53=25%:%
%:%54=26%:%
%:%55=27%:%
%:%56=28%:%
%:%57=29%:%
%:%58=30%:%
%:%59=31%:%
%:%60=32%:%
%:%61=33%:%
%:%62=34%:%
%:%63=35%:%
%:%64=36%:%
%:%65=37%:%
%:%66=38%:%
%:%67=39%:%
%:%68=40%:%
%:%69=41%:%
%:%70=42%:%
%:%71=43%:%
%:%72=44%:%
%:%73=45%:%
%:%74=46%:%
%:%75=47%:%
%:%76=48%:%
%:%77=49%:%
%:%78=50%:%
%:%79=51%:%
%:%80=52%:%
%:%81=53%:%
%:%82=54%:%
%:%83=55%:%
%:%84=56%:%
%:%85=57%:%
%:%86=58%:%
%:%87=59%:%
%:%88=60%:%
%:%89=61%:%
%:%90=62%:%
%:%91=63%:%
%:%92=64%:%
%:%93=65%:%
%:%94=66%:%
%:%95=67%:%
%:%96=68%:%
%:%97=69%:%
%:%98=70%:%
%:%99=71%:%
%:%100=72%:%
%:%101=73%:%
%:%102=74%:%
%:%103=75%:%
%:%104=76%:%
%:%105=77%:%
%:%106=78%:%
%:%107=79%:%
%:%108=80%:%
%:%109=81%:%
%:%110=82%:%
%:%111=83%:%
%:%112=84%:%
%:%113=85%:%
%:%114=86%:%
%:%115=87%:%
%:%116=88%:%
%:%120=91%:%
%:%121=92%:%
%:%122=93%:%
%:%123=94%:%
%:%124=95%:%
%:%125=96%:%
%:%126=97%:%
%:%127=98%:%
%:%128=99%:%
%:%129=100%:%
%:%130=101%:%
%:%131=102%:%
%:%132=103%:%
%:%133=104%:%
%:%134=105%:%
%:%135=106%:%
%:%136=107%:%
%:%137=108%:%
%:%138=109%:%
%:%139=110%:%
%:%140=111%:%
%:%141=112%:%
%:%142=113%:%
%:%143=114%:%
%:%144=115%:%
%:%145=116%:%
%:%149=119%:%
%:%150=120%:%
%:%151=121%:%
%:%152=122%:%
%:%153=123%:%
%:%154=124%:%
%:%155=125%:%
%:%156=126%:%
%:%157=127%:%
%:%158=128%:%
%:%159=129%:%
%:%160=130%:%
%:%161=131%:%
%:%162=132%:%
%:%163=133%:%
%:%164=134%:%
%:%165=135%:%
%:%166=136%:%
%:%167=137%:%
%:%168=138%:%
%:%169=139%:%
%:%170=140%:%
%:%171=141%:%
%:%172=142%:%
%:%173=143%:%
%:%174=144%:%
%:%175=145%:%
%:%176=146%:%
%:%177=147%:%
%:%178=148%:%
%:%179=149%:%
%:%180=150%:%
%:%181=151%:%
%:%182=152%:%
%:%183=153%:%
%:%184=154%:%
%:%185=155%:%
%:%186=156%:%
%:%187=157%:%
%:%188=158%:%
%:%189=159%:%
%:%190=160%:%
%:%191=161%:%
%:%192=162%:%
%:%193=163%:%
%:%194=164%:%
%:%195=165%:%
%:%196=166%:%
%:%197=167%:%
%:%198=168%:%
%:%199=169%:%
%:%200=170%:%
%:%201=171%:%
%:%202=172%:%
%:%203=173%:%
%:%204=174%:%
%:%205=175%:%
%:%206=176%:%
%:%207=177%:%
%:%208=178%:%
%:%209=179%:%
%:%210=180%:%
%:%211=181%:%
%:%212=182%:%
%:%213=183%:%
%:%214=184%:%
%:%215=185%:%
%:%216=186%:%
%:%217=187%:%
%:%218=188%:%
%:%219=189%:%
%:%220=190%:%
